Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-06-23 analysis/TAS_memo/memo_TAS_reviewers.Rmd
updated TAS reviewer memo, paper, references

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    latex_engine: pdflatex
    template: "svm-latex-memo.tex"
    number_sections: true
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
header-includes:
   - \linespread{1.05}

from: Authors
to: The American Statistican
subject: Reply to Reviewer Comments
date: "`r format(Sys.Date(), '%B %d, %Y')`"
memorandum: true
graphics: false
width: .3
---

This memorandum replies to comments provided by the TAS associate editor and two reviewers. We believe the revised paper is much better as a result.

@@

1. As our revised Introduction now makes clear (p. 4-5 and Fig. 1), neither our statistical approach nor any other can alone detect _Batson_ violations, because _Batson_ procedure permits any type of admissible evidence of illegal strike bias -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists. For example, _Batson_ procedure requires the striking attorney at a certain stage to justify the challenged strikes based on (non-discriminatory) reasons. If the attorney fails to produce such reasons, the trial judge must find a _Batson_ violation. If the attorney offers such reasons, the trial judge must assess evidence as to the credibility of those reasons. Our statistical approach cannot adjust the posterior distribution of the bias parameter to account for the credibility of those reasons or other types of evidence (other than the attorney's pattern of strikes) of illegal strike bias. Rather, our approach helps attorneys evaluate the inference of illegal bias in the present case based on historical strike data but _before_ accounting for other evidence relevant to illegal strike bias. The phrase "bias detection" in the paper means only that the credible interval for the bias parameter excludes zero, given the historical strike data, the model, and choice of $\alpha$ value of the power prior. For that "detection" task, our approach lets one assign at most equal weight to past strikes and their strikes in the present case ($\alpha = 1$). In turn, attorneys can use the resulting bias estimate, combined with other evidence, to (1) decide whether to argue _Batson_, (2) to prove the _Batson_ prima facie case, and (3) to persuade the trial judge to find  a _Batson_ violation (i.e., illegal strike bias is more likely than not, given all the evidence).

2. Possible incompatibility of $b_{hist}$ and $b_{curr}$ is always an issue when using past strike data to infer strike bias in the current trial. This is the case for _any_ statistical approach that uses past strike data to infer strike bias in the current trial. And in _Batson_ challenges involving past strike data, the lawyers are already and always arguing about how much weight to assign to that past strike data, albeit with words. By using the power prior, our approach simply makes the weighting of historical strike data more explicit and transparent as a choice of a value for $\alpha$. The more sensitive bias "detection" (i.e. whether the posterior estimate of $b_{curr}$ excludes zero) is to the choice of $\alpha$, the more one should worry about incompatibility. Thus, small vs. large values of $\alpha$ should vary with the suspected degree of incompatibility. In reality, because $b_{curr}$ and $b_{hist}$ are not directly observable, one will need to rely on background knowledge or assumptions about the degree of incompatibility to decide which value of $\alpha$ to choose. In a _Batson_ challenge, the lawyers must disclose and justify their choice of $\alpha$ accordingly. We have revised the paper's Introduction and Discussion (sec 3.1) to emphasize this.

3. In the adversarial setting of the courtroom, our approach thwarts the strategic selection of $\alpha$ values, because it makes the choice of $\alpha$ transparent, and thus the choice of how much to weigh the historical strike data.  As a result, every lawyer will know that any strategic selection of $\alpha$ will be met with scrutiny from the lawyers on the other side. Absent this transparency, lawyers will continue to argue strategically about how much to weigh historical strike data, except with the opacity of words.  We have revised the paper to add this to the Discussion section (sec. 3.1).

4. Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$ in this context of strike bias and _Batson_ challenges. This is because the proper weight to place on historical data depends on what causes incompatibility between the historical data and the current data. The plausible reasons for non-random missingness and such confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things. As we now note in the Discussion section of the paper, measuring incompatibility and choosing $\alpha$ in the face of incompatibility is an active area of research with no clear solution.

5. Choosing a credible interval amounts to deciding how conservative one wishes to be about inferring bias from past strike data, because we define bias "detection" narrowly to mean when the posterior for the bias parameter falls outside the chosen credible interval. Thus, the larger the credible interval one chooses, the greater the chance that zero falls within it. For clarity, we now focus on one credible interval (95%) in the paper and have placed simulation results based on other credible intervals in the Appendix.

6. When we first submitted the paper, we offered to make the app prototype available to the reviewers (as an R package), so they could follow along with section 2.5 of the paper. We received no response to this offer. The offer remains.

@@

@@

It is true, based on our simulations, that bias can be erroneously detected (i.e. even if $b_curr$ = 0), if there are many past trials (e.g., n = 3) and in the bias in those trials exhibits strong incompatibility. On the other hand, given high compatibility, low values of $\alpha$ increase Type I error.  This is why we recommend using different values of $\alpha$ to check for sensitivity. See Item 1(2) above.

Part of the concern here may turn on an error about the standard of proof, as evinced by the use of the word "guilty" in this comment.  _Batson_ violations must be proven under the "preponderance of the evidence" standard (more likely than not), _not_ the more stringent "reasonable doubt" standard that applies when a prosecutor aims to prove a defendant "guilty" of committing the criminal offense charged.

@@

We revised to remove such confusion. On the choice of credible intervals, see Item 1(5) above. The revision does not necessarily fix the type I error, because this coverage probability refers to the true bias parameter (that is fixed in our simulations), whereas we are interested in "bias detection", i.e. whether the credible interval includes 0 or not. The paper also has a sensitivity analysis for the selection of the nominal credible interval level.

@@

This comment is not clear to us. The real data analyses are based on real historical strike data from real criminal cases in the federal district court in Connecticut. If they wish, users of the app can run the model excluding any historical data (as this reviewer suggested) by not selecting the name of a prosecutor or defense attorney. Like our model, the app uses the $N(0,2)$ prior for the bias parameter. Similarly, the app user can alter the weight the historical strike data (if selected) by adjusting $\alpha$ (the power prior parameter). Thus, the app user can: (1) choose not to use historical data, (2) use historical data with equal weight as the current trial (alpha = 1), or (3) down-weight the historical trial data (alpha = 0.5 or alpha = 0.2). See also Item 1(6).

@@

We respectfully disagree with the analogy to "p-hacking". The phrase "p-hacking" typically refers to the practice of using different statistical models on the same data in search of one or more associations that qualify as "statistically significant" (p < .05), and then reporting only the analysis with the model(s) that produce those results. In contrast, in our approach, the statistical model has to qualify as a valid model of the strike procedure of the court of interest. That strike procedure can and must be known and modeled before any strike data is collected or analyzed. More generally, our approach thwarts the strategic selection of $\alpha$ values by making the choice of $\alpha$ transparent. See Item 1(4) and section 3.2 of the revised Discussion.

@@

We agree that, given high incompatibility, (1) the bias "detection" rate can be as high as 100% when current bias is low ($b_curr$); and (2) bias detection can be low where current bias is high but bias in historical trials is low. The tradeoff is that, given compatibility, historical strike data increases power that is otherwise low due the limited number of strikes per trial. The advantage of using the power prior: We can adjust the weight of the historical strike data (and therefore type I error) by adjusting $\alpha$.

@@

This information has been added to the text (sec. 2.5, p. 18).

<!-- A normal distribution was assumed for the distribution of b. The length of the Markov chain was 110000 and the first 10000 iterations were dropped as burn-in. No thinning was performed, as the correlation was weak and the convergence occurred rapidly. Convergence was checked from both traceplots and the Gelman-Rubin statistic. Computation times varies from 0.76 to 1.2 seconds, depending in the amount of data used (current data only _versus_ current and historical data).-->

@@

See Items 1(2) and 1(4) above and section 3.1. of the revised Discussion.

@@

We appreciate this suggestion but we are worried that this would unduly increase the paper's length. Because of this concern, we have already moved other parts of the paper to the (online) Appendix.

@@

Our statistical approach's validity turns neither on how many courts make available the kind of data it requires nor on proving how likely courts will start collecting such data. Indeed, as section 3.2 (paragraphs 1 and 2) of the paper explains, we hope our statistical approach can inform data collection efforts going forward both by courts and courtroom observers exercising their right at attend jury selection in criminal court proceedings.

@@

We assumed no strike bias for the defense attorney only to simplify exposition of the simulation results. Our model of the strike procedure (Equation 1) assumes that the probability of prosecutor strike bias is independent of the probability of defense attorney strike bias. And that model explicitly accounts for the racial composition of the pool of prospective jurors at each strike opportunity.

@@

Thank you for this comment!  The paper now addresses this issue directly in the Discussion, section 3.1, paragraph 4.

@@

No, this is not correct. We have revised the paper to avoid such confusion. See Item 1(5) above. We agree that the choice of credible interval should be consistent within different stages of the _Batson_ challenge and across judges deciding _Batson_ challenges within the same jurisdiction. [NOTE: Simulations use 95% credible interval, while app uses 80% credible intervals.]

@@

Thank you for this reference. The paper now cites to the model in Gastwirth (2005). On "data from an actual case," section 2.5 of the paper (our discussion about the prototype software application) expressly relies on actual strike data from a sample of criminal cases.

@@

On the choice of $\alpha$, see Items 1(2) and 1(3) above and section 3.1 of the revised paper. At most, the historical strike data and the current strike data have _equal_ weight (when $\alpha$ = 1).

@@

Even if it is "[m]ore common" for prosecutors to strike racial minorities, we intend the hypothetical data in Table 8, however "unrealistic", to underscore that, as  _Batson_ law requires, our approach makes no different assumptions for prosecutors than for defense attorneys when estimating attorney strike bias.

>>>>

---
output: 
  pdf_document:
    citation_package: biblatex
    keep_tex: false
    latex_engine: pdflatex
    template: svm-latex-memo.tex
    number_sections: true
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
header-includes:
   - \linespread{1.05}

from: Authors
to: The American Statistican
subject: Reply to Reviewer Comments
date: "`r format(Sys.Date(), '%B %d, %Y')`"
memorandum: true
graphics: false
width: .3
bibliography: references.bib
---

This memorandum replies to comments from a TAS associate editor and two reviewers. We believe the revised paper is much better because of their comments.

@@

1. As our revised Introduction now makes clear (p. 4-5 and Figure 1), neither our statistical approach (nor any other) can alone detect _Batson_ violations, because _Batson_ procedure permits any type of admissible evidence of illegal strike bias -- not just the attorney's past strike data -- and requires the trial judge to weigh all such evidence to decide whether illegal strike bias exists.

    a. For example, _Batson_ procedure requires the striking attorney at a certain stage to justify the challenged strikes based on (non-discriminatory) reasons. If the attorney fails to produce such reasons, the trial judge must find a _Batson_ violation. If the attorney offers such reasons, the trial judge must assess evidence as to the credibility of those reasons. Our statistical approach cannot adjust the posterior distribution of the bias parameter to account for the credibility of those reasons or other types of evidence (other than the attorney's pattern of strikes) of illegal strike bias. Rather, our approach helps attorneys evaluate the inference of illegal bias in the present case based on historical strike data but _before_ accounting for other evidence relevant to illegal strike bias.

    b. The phrase "bias detection" in the paper means only that the credible interval for the bias parameter excludes zero, given the historical strike data, the model, and choice of $\alpha$ value of the power prior. For that "detection" task, our approach lets one assign at most equal weight to past strikes and their strikes in the present case ($\alpha = 1$). In turn, attorneys can use the resulting bias estimate, combined with other evidence, to (1) decide whether to argue _Batson_, (2) to prove the _Batson_ prima facie case, and (3) to persuade the trial judge to find  a _Batson_ violation (i.e., illegal strike bias is more likely than not, given all the evidence).

2. Possible incompatibility of $b_{hist}$ and $b_{curr}$ is always an issue when using past strike data to infer strike bias in the current trial. This is the case for _any_ statistical approach that uses past strike data to infer strike bias in the current trial. And in _Batson_ challenges involving past strike data, the lawyers are already and always arguing about how much weight to assign to that past strike data, albeit with words. By using the power prior, our approach simply makes the weighting of historical strike data more explicit and transparent as a choice of a value for $\alpha$. The more sensitive bias "detection" (i.e. whether the posterior estimate of $b_{curr}$ excludes zero) is to the choice of $\alpha$, the more one should worry about incompatibility. Thus, small vs. large values of $\alpha$ should vary with the suspected degree of incompatibility. In reality, because $b_{curr}$ and $b_{hist}$ are not directly observable, one will need to rely on background knowledge or assumptions about the degree of incompatibility to decide which value of $\alpha$ to choose. In a _Batson_ challenge, the lawyers must disclose and justify their choice of $\alpha$ accordingly. We have revised the paper's Introduction and Discussion (section 4.1) to emphasize this.

3. In the adversarial setting of the courtroom, our approach thwarts the strategic selection of $\alpha$ values, because it makes the choice of $\alpha$ transparent, and thus the choice of how much to weigh the historical strike data.  As a result, every lawyer will know that any strategic selection of $\alpha$ will be met with scrutiny from the lawyers on the other side. Absent this transparency, lawyers will continue to argue strategically about how much to weigh historical strike data, except with the opacity of words.  We have revised the paper to add this to the Discussion section.

4. Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$ in this context of strike bias and _Batson_ challenges. This is because the proper weight to place on historical data depends on what causes incompatibility between the historical data and the current data, which in turn likely varies by. among other things, strike procedure and courtroom practices. As we now note in the Discussion section of the paper, the literature on the power prior discusses information criteria for guide values of $\alpha$ as well as treating $\alpha$ as random and assigning a prior to it. However, that literature is unclear as to what implies for _Batson_ and peremptory strikes, for at least three reasons. First, unlike models of strike procedure discussed here, that literature has mostly focused on clinical trials and models whose properties are well known (e.g., generalized linear models). In the clinical trial setting, it is easier to measure trial-level (and patient-level) characteristics that may be responsible for incompatibility. Even in these settings, selecting a prior for $\alpha$ is not straightforward. @Neelon2010 warn that using random power parameters (in their case, with a beta prior distribution) tends to attenuate (more down-weighting) the influence of the historical data, and so one might have to use highly informative priors on the power parameter. These authors suggest that practitioners consider whether "it is more appropriate to assign $\alpha$ a fixed value based on expert opinion about the relevance of the historical data to the current analysis." More recent versions of the power prior are beginning to address this issue (e.g., using "commensurate power priors" [@Hobbs2011]). Second, our approach does not attempt to maximize coverage in the traditional sense, but aims only to "detect" bias by determining whether zero falls outside the chosen credible interval for the posterior distribution of the bias parameter. Third, whereas some papers treat incompatibility as itself a parameter [e.g., @Ollier2020], the precise approaches rely on strong assumptions that do not obviously make sense in our context.  Accordingly, we believe the most conservative choice (for now) is to keep $\alpha$ a fixed value and check for sensitivity of results to the choice of $\alpha$ value. Proceeding otherwise would require the work of a completely new paper.

5. Choosing a credible interval amounts to deciding how conservative one wishes to be about inferring bias from past strike data, because we define bias "detection" narrowly to mean when the posterior for the bias parameter falls outside the chosen credible interval. Thus, the larger the credible interval one chooses, the greater the chance that zero falls within it. For clarity, we now focus on one credible interval (95%) in the paper and have placed simulation results based on 90% and 80% credible intervals in the Supplementary Material.

6. When we first submitted the paper, we offered to make the app prototype available to the reviewers (as an R package), so they could follow along with section 3.2 of the paper. That section's primary purpose is to show with real data how our approach can be used in practice, including by lawyers in real time during jury selection. We received no response to this offer. The offer remains.

@@

Fourth, the comment is overinclusive. Most work on the use of historical strike data (a) presupposes that _Batson_ law permits prior strikes as relevant evidence of bias in the current trial; (b) presupposes that any statistical approach therein cannot automatically determine _Batson_ violations alone; and (c) adjusts in some way for incompatibility between historical strike data and strikes from the current trial. As an example, we now point in the paper's Introduction to @Gastwirth2014. Our approach has a distinctive benefit of making incompatibility assumptions transparent via choosing and justifying a fixed value for the power prior's $\alpha$.

@@

It is true, based on our simulations, that bias can be erroneously detected (i.e. even if $b_{curr}$ = 0), if there are many past trials (e.g., n = 3) and in the bias in those trials exhibits strong incompatibility. On the other hand, given high compatibility, low values of $\alpha$ increase Type I error.  This is why we recommend using different values of $\alpha$ to check for sensitivity. See Item 1(2) above.

The use of the word "guilty" in this comment implies that part of the concern here may turn on an error about the applicable standard of proof.  _Batson_ violations must be proven under the "preponderance of the evidence" standard (more likely than not), _not_ the more demanding  "reasonable doubt" standard that applies when a prosecutor aims to prove a defendant "guilty" of committing the elements of a criminal offense.

@@

We revised to remove such confusion. This includes changing the prototype software app to use 95% credible intervals. On the choice of credible intervals, see Item 1(5) above. The revision does not necessarily fix the type I error, because this coverage probability refers to the true bias parameter (that is fixed in our simulations), whereas we are interested in "bias detection", i.e. whether the credible interval includes 0 or not. The paper also has a sensitivity analysis for the selection of the nominal credible interval level.

@@

This comment is not clear to us. In section 3.2, we base the real data analyses on real historical strike data from real criminal cases in the federal district court in Connecticut. On how we collected this data, see the paper's Supplementary Materials. If they wish, users of the app can run the model excluding any historical data (as this reviewer suggested) by not selecting the name of a prosecutor or defense attorney. Like our model, the app uses the $N(0,2)$ prior for the bias parameter. Similarly, the app user can alter the weight of the historical strike data (if selected) by adjusting $\alpha$ (the power prior parameter). Thus, the app user can: (1) choose not to use historical data, (2) use historical data with equal weight as the current trial (alpha = 1), or (3) down-weight the historical trial data (alpha = 0.5 or alpha = 0.2). See also Item 1(6).

Similarly, although we did run simulations for when $\alpha = 0$, Figures 2-4 do not depict the bias detection rate for that scenario, simply because that scenario is equivalent to evaluating bias detection with only strike data from the current trial (no historical strike data) -- a topic of prior papers [e.g. @Kadane2018a].

@@

We respectfully disagree with the analogy to "p-hacking". The phrase "p-hacking" typically refers to the practice of using different statistical models on the same data in search of one or more associations that qualify as "statistically significant" (p < .05), and then reporting only the analysis with the model(s) that produce those results. In contrast, in our approach, the statistical model has to qualify as a valid model of the strike procedure of the court of interest. That strike procedure must be known and modeled before any strike data is analyzed. More generally, our approach thwarts the strategic selection of $\alpha$ values by making the choice of $\alpha$ transparent. See Item 1(4) and section 4.2 of the revised Discussion.

@@

We agree that, given high incompatibility, (1) the bias "detection" rate can be as high as 100% when current bias is low ($b_{curr}$); and (2) bias detection can be low where current bias is high but bias in historical trials is low. The tradeoff is that, given high compatibility, historical strike data increases power that is otherwise low due the limited number of strikes per trial. The advantage of using the power prior: We can adjust the weight of the historical strike data (and therefore type I error) by adjusting $\alpha$.

@@

We added this information to the paper (p. 16-17).

@@

See Items 1(2) and 1(4) above and section 4.1 of the revised Discussion.

@@

We appreciate this suggestion but we are worried that this would unduly increase the paper's length, particularly given the lack of a closed form solution for the posterior distribution even in a toy example. Because of this concern, we have already moved other parts of the paper to the Supplementary Materials. We intend the reader to gain understanding by following along using our prototype software application. We also now refer the reader to a gentle introduction to the power prior in @Viele2014.

@@

Our statistical approach's validity turns neither on how many courts make available the kind of data it requires nor on proving how likely courts will start collecting such data. Some courts do, which is how section 3.2 illustrates the approach with real data. Indeed, as section 4.2 of the paper explains, we hope our statistical approach can inform data collection efforts going forward both by courts and courtroom observers exercising their right to attend jury selection in criminal court proceedings.

@@

We assumed no strike bias for the defense attorney only to simplify exposition of the simulation results. Our model of the strike procedure (Equation 1) assumes that the probability of prosecutor strike bias is independent of the probability of defense attorney strike bias. And that model explicitly accounts for the racial composition of the pool of prospective jurors at each strike opportunity. In any case, we re-ran our simulation and found that, as expected, changing the level of defense attorney bias does not alter the bias detection rates for prosecutors (see Table 1 here).

@@

Thank you for this comment!  The paper now addresses this issue directly in the Discussion, section 4.1 (p. 21-22).

@@

No, this is not correct. We have revised the paper and the software application to avoid such confusion. See Item 1(5) above. We agree that the choice of credible interval should be consistent across different stages of a _Batson_ challenge and across judges deciding _Batson_ challenges within the same jurisdiction.

@@

Thank you for this reference. The paper now cites to the model in Gastwirth (2005). On "data from an actual case," section 3.2 of the paper (our discussion about the prototype software application) expressly relies on actual strike data from a sample of actual criminal cases. As mentioned above, we intend the reader to gain understanding by following along using our prototype software application.

@@

On the choice of $\alpha$, see Items 1(2)-1(4) above and section 4.1 of the revised paper. At most, the historical strike data and the current strike data have _equal_ weight (when $\alpha$ = 1).

@@

Even if it is "[m]ore common" for prosecutors to strike racial minorities, we intend the hypothetical strike tally in Table 1 (p. 19), however "unrealistic", to underscore that, as  _Batson_ law requires, our approach makes no different assumptions for prosecutors than for defense attorneys when estimating attorney strike bias.

### 2023-06-23 analysis/memo2_TAS_reviewers.Rmd
updated TAS reviewer memo, paper, references

We have moved Figure 3 to the supplement.

@@

We have revised the figures accordingly.  To make the labels more readable, we used a lighter color for them.  We also fixed the reference to $b_curr$.

@@

See below.

## Associate Editor

@@

We have now provided links to the app as well as the code used to generate the simulation results. Instructions for using the app are now provided not only in section 3.2 of the paper and the Supplemental Materials, but also a vignette in the app itself [TBD].

@@

We have revised the paper in paragraph 2 to expressly distinguish between "bias" as the illegal strike bias (the parameter to be estimated) from "bias" in the statistical sense.  We hope this reduces any confusion.[EDITME]

@@

We now introduce the term "cognizable class" later, in section 2. By "cognizable" class, we mean the social category to which the struck juror belongs that is the possible basis of illegal strike bias. For example, if the focus is on strike bias against Black prospective jurors because of their race, the "cognizable" class is "Black prospective jurors". If, however, the focus is on strike bias against Black female prospective jurors because of their race and gender, the "cognizable" class is "Black female prospective jurors."  Because _Batson_ applies without regard to the "majority" or "marginalized" status of struck jurors (e.g., Black prospective jurors _and_ White prospective jurors), we have avoided those terms.

We use the phrase "cognizable class members subject to strike" instead of "cognizable class jurors", for two reasons. First, at the time any strike occurs, no one is a "juror", because no one has yet been seated onto the jury. Second, at the time any strike occurs, not every prospective juror could have been struck. As we assume in the paper, only a subset of prospective jurors present during jury selection are ever subject to strike at any one time. The size of that subset is usually based on the number of seats of the jury itself plus the total number of strikes that the parties could exercise.

@@

_(page 9, lines 3-19) I’m still a little unsure exactly how bhist and bcurr are used to generate the data. I’m assuming you’re generating it via the probability model described in Equation (2) given those selected values, is that right?_

@@

_(page 10, lines 25-29) Does the application allow users to revise the historical data used? That is, should there be guidance about choosing between low $\alpha$ with all the historical data or large α with a subset of the historical data? That is, it seems your approach can be compatible with the stratified approach of Gastwirth and Xu (2014) – would that improve the issue with heterogeneous historical data?_

Subsetting the historical data does not necessarily reduce the degree of incompatibility between the historical strike data and the current trial ($b_hist$ - $b_curr$). As we discuss in the paper, it depends on what one assumes about how much the subsetting criteria (e.g., the salient criteria for "similar" cases) changes the degree of incompatibility. We stress "assumes", because the true values of $b_hist$ and $b_curr$ are not directly observable in actual cases. Accordingly, for any subset of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive the posterior estimate of illegal strike bias is to different values of $\alpha$. The current version of the software prototype does not let users subset the historical data by some characteristic (e.g., date, defendant race). However, in principle, it could be modified to add that kind of feature.

## Reviewer 2

@@

Thank you for this comment.  We have revised the paper accordingly. [TBD]

@@

Thank you for this comment.  We have revised the Discussion to discuss this further. [TBD].

We take this comment to posit that a judge is unlikely to find ("seriously entertain") even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal bias against the cognizable class (e.g. Black prospective jurors), because $b = 0$ falls within the credible interval, even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal bias, i.e., $b = 0$ falls outside the credible interval.  This might occur, for example, where an attorney is motivated to strike Black prospective jurors because they are Black _only_ in cases with Black defendants; all that attorney's past trials involved Black defendants; but the current trial has a White defendant. In this extreme scenario, we agree that a judge might properly worry about high incompatibility and not find a _Batson_ prima facie case for that reason, even though no one can directly observe or know with complete certainty whether this is a case where $b_curr = 0$ and $b_hist >> 0$. Or the judge may still find enough for a prima face case, but downweight the historical data when ultimately deciding whether illegal bias likely occurred. In the reverse extreme situation, the posterior estimate of $b$ with $\alpha = 0$ detects illegal bias but does not detect illegal bias when $\alpha > 0$. Here, too, we agree that a judge might properly worry about high incompatibility, except that down-weighting the historical data here increases the support for finding at least enough a _Batson_ prima facie case.

>>>>

We have moved what was previously Figure 3 to the supplement.

@@

We have revised the figures accordingly.  To make the labels more readable, we used a lighter color for them.  We also fixed the reference to $b_{curr}$.

@@

Section 3.2 of the paper and the related Supplementary Material illustrates our approach using our prototype software application and a _real_ data set of historical strike data. For details on this dataset, see section 2.1 of the Supplementary Material. Given our likelihood function (see Equation (3)), there is no closed-form solution for the posterior distribution of $b$. This is why we used our prototype software application to show how the posterior estimate of $b$ changes (see Figures 7 and S8).

We have made the app available to the public as a R package that includes the underlying dataset of historical strike data. We hope this will help statistics instructors in the way the TAS Associate Editor has in mind. However, because of time constraints, we are simply unable to provide in the paper's Supplement a complete tutorial tailored for classroom instruction by teachers of Bayesian statistics. On an introduction to a Bayesian approach to estimating illegal strike bias (albeit without historical strike data), we recommend @Kadane2018a and @Kadane2018b.

# Associate Editor

@@

The paper now contains links to the app as well as the code we used to generate the simulation results. We intend section 3.2 of the paper (and related Supplementary Material) as instructions for using the app. Those links are identifying, which is why we omitted them from the last draft of the paper.

@@

We appreciate this suggestion, but we worry that making this point here in terms of Type I and Type II error might lead to net confusion. The statement that "errors of Type I are okay" is fine if it refers to the idea that at step two of the _Batson_ sequence (whether to find a _prima facie_ case), one should tolerate more Type I error (falsely inferring illegal strike bias), because all it does is lead the judge to force the striking attorney to offer up reasons for the strikes and then have the judge assess the credibility of those reasons along with the other evidence. Type I errors, however, are far less tolerable -- not as okay -- at step four, where the judge has to consider historical strike data along with all the other evidence and ultimately find whether illegal strike bias is more likely than not. And Type I errors are worse still as the judge assigns more relative weight to any inference from historical strike data as compared to other items of evidence.

@@

We have revised the paper to introduce the term "cognizable class" in section 2 instead of the Introduction. By "cognizable class", we mean the social category to which the struck juror belongs that the law accepts as a possible basis of illegal strike bias. For example, if the focus is on strike bias against Black prospective jurors because of their race, the cognizable class is "Black" prospective jurors. If, however, the focus is on strike bias against Black female prospective jurors because of their race and gender, the cognizable class is "Black female" prospective jurors. And if the focus is on illegal strike bias against White prospective jurors because of their race, the cognizable class is "White" prospective jurors. Because _Batson_ law applies without regard to the "majority" or "marginalized" status of prospective jurors, we have avoided such terms.

We use the phrase "cognizable class members subject to strike" instead of "cognizable class jurors" for two reasons. First, at the time any strike occurs, no one is a "juror", because no one has yet been seated onto the jury. Second, at the time any strike occurs, not every prospective juror could have been struck. As we assume in the paper, only a subset of prospective jurors present during jury selection are ever subject to strike at any one time. Although courts vary in practice, the size of that subset is often based on the number of seats of the jury itself plus the total number of strikes that the parties could exercise.

@@

_(page 9, lines 3-19) I’m still a little unsure exactly how b_hist and b_curr are used to generate the data. I’m assuming you’re generating it via the probability model described in Equation (2) given those selected values, is that right?_

@@

_(page 10, lines 25-29) Does the application allow users to revise the historical data used? That is, should there be guidance about choosing between low $\alpha$ with all the historical data or large $\alpha$ with a subset of the historical data? That is, it seems your approach can be compatible with the stratified approach of Gastwirth and Xu (2014) – would that improve the issue with heterogeneous historical data?_

Subsetting the historical data strike does not necessarily reduce incompatibility between the historical strike data and the current trial. Rather, it depends on what one _assumes_ about how much the subsetting criteria (e.g., the salient criteria for "similar" cases) affects incompatibility. We stress "assumes", because the true values of $b_{hist}$ and $b_{curr}$ are not directly observable in actual cases. Accordingly, for any _subset_ of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive the posterior estimate of illegal strike bias is to different values of $\alpha$. The current version of the software prototype does not let users subset the historical data by some trial-level characteristic (e.g., defendant race). However, in principle, it could be modified to add that feature.

_(page 14, Figure 5) the figure legend makes it challenging to compare between 1 and 2, though one could anticipate 2 is the dashed line._

We have revised that figure legend to better distinguish between the different line types.

_(page 16, line 53-54) 110000 doesn’t have a comma and 10,000 does._

Thank you for catching this.  We have revised accordingly.

_(page 19, Figure 8) This demonstration is very helpful for seeing the utility of the app. This makes me think of two possible improvements: 1. See point three of the other reviewer. This situation presents the opportunity to discuss any disagreements between current and historical data._

We have revised the Discussion accordingly.  See also our reply to other reviewer below.

_I understand the concern about paper length . . . There is value to the general audience for demonstrating posterior changes as the reviewer discusses. While this would make the paper long (and require too many figures), I do believe this would be helpful supplementary material._

See above reply to TAS editor on this point.

# Reviewer 2

@@

Thank you for suggesting this contrast between _Batson_ law and these State laws. We have revised the paper to mention on page 1 that for _Batson_ challenges, the "ultimate inquiry" is whether the striking party "was motivated in substantial part by discriminatory intent." (@flowers2019, p. 2241) We did not use the phrase "significant factor," because it does not appear in _Miller-El v. Dretke_, 545 U.S. 231 (2005). At best, the Court there wrote: "Comparing his strike with the treatment of panel members who expressed similar views supports a conclusion that race was significant in determining who was challenged and who was not. _Id._ at 252 (footnote omitted).

@@

We took this comment to have two intended meanings.

First, we read this comment to posit that a judge is unlikely to find even a _Batson_ prima facie case where "data in the present case indicated the removal rates of protected and unprotected members of the venire were (essentially) equal." However, in some cases, illegal strike bias can produce equal strike _rates_.  To illustrate, consider the following hypothetical pattern of strikes by a prosecutor in a trial with a strike procedure like the one we model in the paper.

```{r}
#| label: hypodata
#| echo: false

round <- c(1:10)
num_cog <- c(7,6,5,4,3,2,2,2,2,2)
total <- c(30,28,26,24,22,20,18,16,14,12)
cog <- c(1,1,1,1,1,0,0,0,0,0)

df <- data.frame(round, num_cog, total, cog)

knitr::kable(df, caption = "Hypothetical Strikes by Prosecutor")

```

Here, the strike rates for the cognizable class and the non-cognizable class are equal (5 out of 10). However, using the paper's approach and assuming no historical strike data (or $\alpha = 0$), we detect illegal strike bias, i.e., the 95% credible interval for the posterior estimate of $b$ excludes zero.  The reason turns on the relatively few members of the cognizable class that could have struck in each round. Thus, in this hypothetical case, the judge would err by relying on equal strike rates to infer no illegal strike bias.

Second, we took this comment to say that a judge is unlikely to find even a _Batson_ prima facie case if the posterior estimate of $b$ with $\alpha = 0$ detects no illegal strike bias (the credible interval includes zero), even if the posterior estimate of $b$ given some set of non-zero values of $\alpha$ detects illegal strike bias (the credible interval excludes zero). 
We have revised the Discussion to further address high incompatibility. As we observe there, because no one can directly observe $b_{curr}$ and $b_{hist}$, they cannot know for sure whether their case involves high incompatibility. Nonetheless, if the posterior estimate of $b$ is highly sensitive to the choice of $alpha$ -- something one can observe -- then we agree that a judge might properly worry about erroneously detecting strike bias due to high incompatibility. As for what a judge might do in this situation, a judge might not find a _Batson_ prima facie case for that reason. Or the judge may still find enough for a prima face case, but down-weight the historical data (by selecting an $\alpha$ value less than 1) when ultimately deciding whether illegal strike bias likely occurred. In the reverse situation of high incompatibility, the posterior estimate of $b$ with $\alpha = 0$ detects illegal bias but (erroneously) does not detect illegal bias when $\alpha > 0$. Here, we agree that a judge should worry about erroneously failing to detect strike bias due to high incompatibility. Down-weighting the historical data here increases the support for a finding that illegal strike bias likely occurred.
